---
title: "final_gov_project"
author: "Shai-Li Ron"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(janitor)
library(primer.data)
library(rstanarm)
library(readxl)
library(lubridate)
library(ggthemes)
library(patchwork)
library(ggplot2)
library(hrbrthemes)
library(ggdist)

```


# Method:
This project examines the Covid 19 situation in Israel, a world leading nation in the percentage of vaccinated individuals.
We are going to ask the following questions:
1) what is the relationship between the percentage of vaccinated individuals per city and the number of active cases?
2) What is the relationship between socio-economic status of a city and vaccination rate?
3) what is the relationship between minority populations and vaccination rates?


# Percentage vaccinate vs active cases - total Israel
```{r}
# First I loaded new case data, by date. 
new_cases <- read_excel(path = "Raw_data/new_covid_cases.xlsx") 

# Next I loaded vaccinated data, by date. 
total_vac <- read_excel(path = "Raw_data/number_vaccinated.xlsx") 
```

```{r}
# I then needed to join the two data sets. I plan to do this by date. I chose to
# begin with the new_case data as that was happening before the vaccine rates.

total_case_vax <- new_cases %>% 
  left_join(total_vac, by = "date") %>% 

# It makes sense that we are getting NA values, as vaccination only started a
# long time after the cases started.
  
  select(date, new_cases, first_dose, second_dose) %>% 
  separate(col = date,
          into = c("day", "month", "year"),
          sep = "-") %>% 
  mutate(date = make_date(year, month, day)) %>% 
  select(date, new_cases, first_dose, second_dose)  

# First I needed to separate the original date type I had (EU, Israel type),
# then I used make date from the lubridate library to make a proper date that R
# can use.

vax_plot <- total_case_vax %>% 
  # filter(date > as.Date("2020-12-01")) %>% 
     pivot_longer(cols = c(first_dose, second_dose), 
               names_to = "type", values_to = "value") %>% 
  ggplot(aes(x = date,
             y = value,
             color = type)) + 
  geom_line() +
  theme_fivethirtyeight() +
  scale_y_continuous(labels = scales::number_format(accuracy = 100)) +
  labs(title = "Number of Israelis Vaccinated: First and Second Dose",
       x = "Date",
       y = "Number of Vaccinatated") +
  geom_label(x=as.Date("2020-12-19"), y = -100, 
             label = "First Vaccines", color = "black", size = 2) +
  geom_label(x=as.Date("2021-03-19"), y = 4526000, 
             label = "50% of Population Vaccinated", color = "black", size = 2) +
  scale_color_discrete(name = " ", labels = c("First Dose", "Second Dose"))
  
# FIX Y-axis  
# Why are my x and y axis not showing?
```


```{r}
new_case_plot <- total_case_vax %>% 
  ggplot(aes(x = date,
             y = new_cases)) + 
  geom_line() +
  labs(title = "Number of New Covid Cases",
       x = "Date",
       y = "Number of Vaccinatated") +
  theme_fivethirtyeight() 

```


```{r}
new_case_plot / vax_plot +
  plot_annotation(title = "Number of New Covid Cases and Vaccinated in Israel",
                  subtitle = "New Covid Cases Decreased when Vaccination Rate Increased",
                  caption = "Source: Israel Ministry of Health") &
  coord_cartesian(ylim = c(0, .035))


```

```{r}
# Trying to plot this with two different axis

coeff <- 100
temperatureColor <- "#69b3a2"
priceColor <- rgb(0.2, 0.6, 0.9, 1)
black <- "#000000"

total_case_vax %>% 
     pivot_longer(cols = c(first_dose, second_dose), 
               names_to = "type", values_to = "vax") %>% 
  filter(date > as.Date("2020-06-01")) %>% 
  ggplot(aes(x = date, color = type)) +
  geom_line(aes(y=vax/coeff), size=1) + 
  geom_line( aes(y=new_cases), size=1, color=priceColor) +
  
  scale_y_continuous(
    
    # Features of the first axis
    name = "New Covid Cases",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Number of Vaccinated Israelis")
  ) + 
  
  theme_ipsum() +

  theme(
    axis.title.y = element_text(color = priceColor, size=10),
    axis.title.y.right = element_text(color = black, size=10)
  ) +

  labs(title = "New Covid Cases and Vaccination in Israel",
       subtitle = "New Covid Cases Decreased when Vaccination Rate Increased",
       caption = "Source: Israel Ministry of Health",
       x = "Date") +
  scale_color_discrete(name = " ", labels = c("First Dose", "Second Dose")) +
  geom_label(x=as.Date("2020-12-19"), y = -10, 
             label = "First Vaccines", color = "black", size = 2) +
  geom_label(x=as.Date("2021-03-19"), y = 45260, 
             label = "50% Vaccinated", color = "black", size = 2)
  

# FIX AXIS 
# Try to add color of line in the same as first dose and second dose.
```

# Percetange vaccinated vs. Active cases - by city
```{r}
# The first data set we will look at is the most recent data set for the percentage of vaccinated per city. 
# Loading vaccinated by city
vax_by_city_age <- read_excel(path = "Raw_data/total_bycity.xls") 
perc_vax_city <- read_excel(path = "raw_data/perc_city.xlsx")
city_econ <- read_excel(path = "raw_data/econ_status_city.xlsx")

```

```{r}

vax_by_city_age %>% 
  clean_names() %>% 
  mutate_at(vars(-c("city")), as.numeric) %>%  
  
# Any indefinite value of <15 was ignored - replaced by NA
  
  group_by(city) %>%
  mutate(total = sum(x16_19, x20_29, x30_39, x40_49, x50_59, x60_69, x70_79, x80_89, x90))

```

# active_per_10000 vs. percent vaccinated
```{r}
perc_vax_city %>% 
   mutate_at(vars(-c("city")), as.numeric) %>% 
   arrange(perc_first_dose) %>% 

# Had to make these vectors numeric so I could manipulate them.
# We chose to look at second dose - as this is what matters for being fully vaccinated.

  select(city, perc_second_dose, active_per_10000) %>% 
  ggplot(aes(x = perc_second_dose, 
             y = active_per_10000)) +
  geom_point() +
  geom_smooth(method = "lm", 
              formula = y ~ x) +
  labs(title = "Relationship between Active Cases per 10,000 People and Vaccination Percentage In Cities",
       subtitle = "A negative relationship exists between active cases and vaccination percentage",
       x = "Percent Vaccinated Second Dose in Given City",
       y = "Number of Active Cases per 10,000 People in Given City",
       caption = "Source: Israel Health Ministry") +
    theme_gray()


```


```{r}
clean_perc_vax_city <- perc_vax_city %>% 
   mutate_at(vars(-c("city")), as.numeric)  

  
```


```{r}
vax_econ_clean <- city_econ %>% 
  clean_names() %>% 
  group_by(city) %>% 
  full_join(clean_perc_vax_city, by = "city") %>% 
  mutate("rank2017" = as.numeric(cluster_2017_4))

vax_econ_clean

vax_econ_clean %>% 
  select(city, rank2017, city_rank, perc_second_dose, active_per_10000) %>% 
  ggplot(aes(x = rank2017, 
             y = perc_second_dose,
             colour = city_rank)) +
  geom_point() +
  scale_x_continuous(labels = scales::number_format(accuracy = 1), breaks = 0:10) +
  labs(title = "Relationship Between City's Socioeconomic Rank and Vaccination Percentage",
       subtitle = "Cities with a higher socio-economic status have a higher vaccination percentages",
       x = "City Socio-Economic Rank (2017)",
       y = "Percentage Vaccinated (Second Dose)",
       caption = "Source: Israel Health Ministry") +
  geom_smooth(method = "lm", 
              formula = y ~ x) +
  scale_color_gradient(low="blue", high="red") 
                      

```

# Predictive Model 
```{r}

fit_1 <- stan_glm(data = vax_econ_clean, 
                    formula = perc_second_dose ~ rank2017, 
                    family = gaussian, 
                    refresh = 0,
                    seed = 9)
fit_1 %>% 
   as_tibble() %>% 
  rename(mu = `(Intercept)`) %>% 
  ggplot(aes(x = mu)) +
    geom_histogram(aes(y = after_stat(count/sum(count))), 
                   bins = 100) +
    labs(title = "Posterior Distribution for Percentage of Vaccinated",
         subtitle = "The model predicts around 32% vaccinated in any randomly selected city",
         x = "Percent Vaccinated Second Dose",
         y = "Probability",
         caption = "Data source: Israel Health Ministry") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme_classic()

```

```{r}
# What would be the percentage vaccinated in 10 cities of different socio-economic status? 

newobs <- tibble(rank2017 = 1:10)

pp_10 <- posterior_predict(fit_1, newdata = newobs) %>%
    as_tibble() 

pp_10 %>%
  pivot_longer(cols = `1` : `10`,
               names_to = "status",
               values_to = "perc_vax") %>% 

# We need a single row for the status and another one for the perc_vax.

  ggplot(aes(x = perc_vax, y = fact_reorder(status, perc_vax)) +
      stat_halfeye(aes(fill = stat(cut_cdf_qi(cdf, .width = c(0.95, 1)))),
               show.legend = FALSE) +
  scale_fill_calc() +
  labs(title = "Posterior for Percentage Vaccinated in Cities of Different Socio-Economic Status",
       x = "Percent of Vaccinated in City",
       y = "Socio-Economic Status",
       caption = "Source: Israel Heatlh Ministry")
    
```

```{r}
pp_10 %>%
  clean_names() %>% 
  pivot_longer(cols = x1:x10,
               names_to = "status",
               values_to = "perc_vax") %>% 

# We need a single row for the status and another one for the perc_vax.

  ggplot(aes(x = perc_vax, y = status)) +
  stat_slab(position = "dodge") +
  guides(alpha = FALSE) 
```

